---
title: "testShiny"
author: "Bivin"
date: "4/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let the user read in the data (the Beer Data.) 
Let the user make a histogram of the IBU
Let the user make a histogram of the ABV  (separate histogram)
Low the user switch back and forth between histograms and boxplots to represent this information.  You could do this with a radio button that switches the plot from histogram to boxplot … the implementation however is up to you.  
Allow the user to filter the histograms / boxplots above by state. 
In addition to the histograms, add a scatter plot of IBU v. ABV
Add the ability to add or take off the simple linear regression line. 
Allow the user to filter the data on the scatter plot by state.
Add at least one additional plot from your analysis and make sure to clearly label and/or describe the information it is conveying.  



## Rshiny Test
This is a test of an RShiny app in RMarkdown

```{r }
library(shiny)

ui <- fluidPage(
  
  # App title ----
  titlePanel("Brewery Study Interactive"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      
      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
      
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      
      # Output: Histogram ----
      plotOutput(outputId = "distPlot")
      
    )
  )
)


# Define server logic required to draw a histogram ----
server <- function(input, output) {
  library(tidyverse)
library(ggthemes)
library(caret)
library(mvtnorm)
library(class)
library(e1071)
library(usmap)
  beers <- read.csv("Beers.csv",header = TRUE)
  breweries <- read.csv("Breweries.csv",header = TRUE)
  
  # Histogram of the ABV
  output$distPlot <- renderPlot({
   hist(x=, )
    x    <- beers$ABV
    bins <- seq(min(x), max(x), length.out = input$bins + 1)
    
    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Alchol By Volume",
         main="Count of Beers per ABV")
  })
  
}


shinyApp(ui, server)

```

